\TODO{Run proofreader on all text (temporarily disabled because it slowed down IDE)}

We use standard definitions, sourced from the book by \citeauthor{BB98} \citetitle*{BB98}, modified and extended as needed.

\begin{defn}[graph]
  A \emph{graph} $G$ is an ordered pair of disjoint sets $(V, E)$ such that $E$ is the subset of the set $V \choose 2$ that is of unordered pairs of $V$.
\end{defn}

\begin{wrapfigure}{r}{0.4\textwidth}
  \label{fig:examplePath}
  \centering\begin{tikzpicture}[scale=.7, simplegraph]
    \node(a) at (0, 0) {$v_1$};
    \node(b) at (2, 0) {$v_2$};
    \node(c) at (4, 0) {$v_3$};
    \node(d) at (6, 0) {$v_4$};

    \draw(a) to (b);
    \draw(b) to (c);
    \draw(c) to (d);

    \draw[dashed](a) to[in=90, out=90] (d);
    \draw[dashed](a) to[in=90, out=90] (c);
    \draw[dashed](b) to[in=90, out=90] (d);
  \end{tikzpicture}
  \caption{An example graph $G_0$}
\end{wrapfigure}

We will only consider finite graphs, that is $V$ and $E$ are always finite. If $G$ is a graph, then $V = V(G)$ is the \emph{vertex set} of $G$, and $E = E(G)$ is the \emph{edge set}. The size of the vertex set of a graph $G$ will be called the \emph{cardinality} of $G$.

An edge $\{x, y\}$ is said to \emph{join}, or be between vertices $x$ and $y$ and is denoted by $xy$. Thus $xy$ and $yx$ mean the same edge (all our graphs are \emph{unordered}). If $xy \in E(G)$ then $x$ and $y$ are adjacent, connected or neighboring. By $N(x)$ we will denote the \emph{neighborhood} of $x$, that is all vertices $y$ such that $xy$ is an edge. If $xy \notin E(G)$ then $xy$ is a \emph{nonedge} and $x$ and $y$ are \emph{anticonnected}.

Figure \ref{fig:examplePath} shows an example of a graph $G_0 = (V, E)$ with $V = \{v_1, v_2, v_3, v_4\}$ and $E = \{v_1v_2, v_2v_3, v_3v_4\}$. We will mark edges as solid lines on figures. Nonedges significant to the ongoing reasoning will be marked as dashed lines.

\begin{defn}[subgraph]
  $G' = (V', E')$ is a \emph{subgraph} of $G = (V, E)$ if $V' \subseteq V$ and $E' \subseteq E$.
\end{defn}

\begin{defn}[induced subgraph]
  If $G' = (V', E')$ is a subgraph of $G$ and it contains \emph{all edges} of $G$ that join two vertices in $V'$, then $G'$ is said to be \emph{induced subgraph} of $G$ and is denoted $G[V']$.
\end{defn}

Given a graph $G = (V, E)$ and a set set $X \subseteq V$ by $G\setminus X$ we will denote a induced subgraph $G[V\setminus X]$.

For example $(\{v_1, v_2, v_3\}, \{v_1v_2\})$ is \emph{not} an induced subgraph of the example graph $G_0$, while $(\{v_1, v_2, v_3\}, \{v_1v_2, v_2v_3\}) = G_0[\{v_1, v_2, v_3\}] = G_0 \setminus \{v_0\}$ is.

\begin{defn}[$X$-completeness]
  Given set $X \subseteq V$, vertex $v \notin X$ is \emph{$X$-complete} if it is adjacent to every node $x \in X$. A set $Y \subseteq V$ is $X$-complete if $X \cap Y = \emptyset$ and every node $y \in Y$ is $X$-complete.
\end{defn}

\begin{defn}[path]
  A \emph{path} is a graph $P$ of the form
  \[ V(P) = \{x_1, x_2, \ldots, x_l\},\quad E(P) = \{x_1x_2, x_2x_3, \ldots, x_{l-1}x_l\} \]
\end{defn}
This path $P$ is usually denoted by $x_1x_2\ldots x_l$ or $x_1 - x_2 - \ldots - x_l$. The vertices $x_1$ and $x_l$ are the \emph{endvertices} and ${l-1} = |E(P)|$ is the \emph{length} of the path P. $\{x_1, \ldots x_{l-1}\}$ is the \emph{inside} of the path $P$, denoted as $P^*$.

Graph $G_0$ is a path of length 3, with the inside $G_0^* = \{v_2, v_3\}$. If we would add any edge to $G_0$ it would stop being a path (sometimes we call such an edge a \emph{chord}).


\begin{defn}[connected graph, subset]
  A graph $G$ is \emph{connected} if for every pair $\{x, y\} \subseteq V(G)$ of distinct vertices, there is a path from $x$ to $y$. \todo{formal enough?}
  A subset $X \subseteq V(G)$ is connected if the graph $G[X]$ is connected.
\end{defn}

\begin{defn}[component]
  A \emph{component} of a graph $G$ is its maximal connected induced subgraph.
\end{defn}


\begin{defn}[cycle]
  A \emph{cycle} is a graph $C$ of the form
  \[ V(C) = \{x_1, x_2, \ldots, x_l\},\quad E(C) = \{x_1x_2, x_2x_3, \ldots, x_{l-1}x_l, x_lx_1\} \]
\end{defn}

This cycle $C$ is usually denoted by $x_1x_2\ldots x_lx_1$ or $x_1 - x_2 - \ldots - x_l - x_1$. $l = |E(C)|$ is the \emph{length} of the cycle $C$. Sometimes we will denote the cycle of length $l$ as $C_l$.

Notice, that a cycle is not a path (nor is a path a cycle). If we add an edge $v_1v_4$ to the path $G_0$ it becomes an even cycle $C_4$.

\begin{defn}[hole]
  A \emph{hole} is a cycle of length at least four.
\end{defn}

If a path, a cycle or a hole has an odd length, it will be called \emph{odd}. Otherwise, it will be called \emph{even}.

\begin{defn}[complement]
  A \emph{complement} of a graph $G = (V, E)$ is a graph $\overline{G} = (V, {V \choose 2} \setminus E)$, that is two vertices $x, y$ are adjacent in $\overline{G}$ iff they are not adjacent in $G$.
\end{defn}

% Sometimes, we will call a complement of a member of a class $\Gamma$ an \emph{anti-
% $\Gamma$}, e.g. graph $G = (\{v_1, v_2, v_3, v_4\}, \{v_1v_3, v_2v_4\})$ is an anticycle.

\begin{defn}[anticonnected graph, subset]
  A graph $G$ is anticonnected if $\overline{G}$ is connected.
  A subset $X$ is anticonnected if $\overline{G}[X]$ is connected.
\end{defn}

\begin{defn}[anticomponent]
  An \emph{anticomponent} of a graph $G$ is an induced subgraph whose complement is a component in $G$.
\end{defn}

\begin{defn}[clique]
  A \emph{complete} graph or a \emph{clique} is a graph of the form $G = (V, {V \choose 2})$, that is every two vertices are connected. We will denote a clique of $n$ vertices as $K_n$.
\end{defn}

\begin{defn}[clique number]
  A \emph{clique number} of a graph $G$, denoted as $\omega(G)$, is a cardinality of its largest induced clique.
\end{defn}

\begin{defn}[anticlique]
  An \emph{anticlique} is a graph in which there are no edges. We will also call anticliques \emph{independent sets}. 
\end{defn}
In a similar fashion, given a graph $G = (V, E)$, a subset of its vertices $V' \subseteq V$ will be called \emph{independent} (in the context of $G$) iff $G[V']$ is an anticlique.

\begin{defn}[stability number]
  A \emph{stability number} of a graph $G$, denoted as $\alpha(G)$, is a cardinality of its largest induced stable set.
\end{defn}

\begin{defn}[coloring]
  Given a graph $G$, its \emph{coloring} is a function $c: V(G) \rightarrow \mathbb{N}^+$, such that for every edge $xy \in E(G)$, $c(x)$ is different from $c(y)$. A $k-coloring$ of $G$ (if exists) is a coloring, such that for all vertices $x \in V(G)$, $c(x) \leq k$.
\end{defn}

\begin{defn}[chromatic number]
  A \emph{chromatic number} of a graph $G$, denoted as $\chi(G)$, is a smallest natural number $k$, for which there exists a $k$-coloring of $G$.
\end{defn}

\begin{defn}[Berge graph]
  \label{def:bergeGraph}
  A graph $G$ is \emph{Berge} if both $G$ and $\overline{G}$ have no odd hole.
\end{defn}

\begin{defn}[perfect graph]
  \label{def:perfectGraph}
  A graph $G$ is \emph{perfect} if for its every induced subgraph $G'$ we have $\chi(G') = \omega(G')$
\end{defn}

\begin{defn}[C-major vertices]
  Given a shortest off hole $C$ in $G$, a node $v \in V(G) \setminus V(C)$ is $C-major$ if the set of its neighbors in $C$ is not contained in any 3-node path of $C$.
  \TODO{a picture of this, clean odd hole, amenable hole}
\end{defn}

\begin{defn}[clean odd hole]
  An odd hole $C$ of $G$ is \emph{clean} if no vertex in $G$ is $C$-major.
\end{defn}

\begin{defn}[cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{cleaner for $C$} if $X \cap V(C) = \emptyset$ and every $C$-major vertex belongs to $X$.
\end{defn}

Let us notice, that if $X$ is a cleaner for $C$ then $C$ is a clean hole in $G \setminus X$.

\begin{defn}[near-cleaner]
  Given a shortest odd hole $C$ in $G$, a subset $X \subseteq V(G)$ is a \emph{near-cleaner for $C$} if $X$ contains all $C$-major vertices and $X \cap V(C)$ is a subset of vertex set of some 3-node path of $C$.
\end{defn}

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if it is a shortest off hole in $G$, it is of length at least 7 and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}